Stochastic Variance Reduced Policy Gradient

### 2018-01-29 svrpg.tex
added comments to main theorem proof

%%Theorems
\usepackage{amsthm}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{conj}{Conjecture}
\newtheorem{example}{Example}
%Restatable
\usepackage{thmtools, thm-restate}
\declaretheorem[numberwithin=section]{thm}
\declaretheorem[sibling=thm]{lemma}
\declaretheorem[sibling=thm]{corollary}
\declaretheorem[sibling=thm]{assumption}
\declaretheorem[sibling=thm]{theorem}
%e.g. ...
\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
% graphics
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
%% TODOs
% \setlength{\marginparwidth}{2cm}
\usepackage[colorinlistoftodos, textwidth=20mm]{todonotes}
\definecolor{citrine}{rgb}{0.89, 0.82, 0.04}
\definecolor{blued}{RGB}{70,197,221}
%%todo by Marcello
\newcommand{\todomarc}[1]{\todo[color=green, inline]{\small #1}}
\newcommand{\todomarcout}[1]{\todo[color=green]{\scriptsize #1}}
%%todo by Matteo Papini
\newcommand{\todomat}[1]{\todo[color=citrine, inline]{\small #1}}
\newcommand{\todomatout}[1]{\todo[color=citrine]{\scriptsize #1}}
%%todo by Matteo Pirotta
\newcommand{\todopir}[1]{\todo[color=blued, inline]{\small #1}}
\newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
%%%%%%

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
\newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla J(#1)}
\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%Specific
\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\mid s}\left[#2\right]}
\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\mid s}\left(#2,#3\right)}
\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\mid s}\left[#2\right]}
\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
\newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
%%%%%%

@@

\section{Algorithm}
In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment. We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{itemize}
	\item The objective $J(\vtheta)$ is typically non-convex;
	\item The dataset $\mathcal{T}$ can be infinite;
	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
\end{itemize}
To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?]. Because of the infinite dataset, we can only rely on an estimate of the full gradient. Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories. We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward. The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
\begin{align*}
	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
\end{align*}
where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. We immediately have the following (proof in Appendix \ref{app:proofs}):

\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
The SVRPG gradient estimator is unbiased:
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
\end{restatable}

[May also add Lemma about convergence of the variance to zero, since it provides an intuitive justification of some design choices]
The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient and the step size $\alpha$. Adaptive selection of these meta-parameters is also possible. However, the next section will focus on providing converge guarantees for the plain SVRPG algorithm.

@@

\section{Convergence Guarantees}
\begin{assumption}\label{ass:bounded_score}
For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
\begin{align*}
	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq G \\
	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq F
\end{align*}
\end{assumption}

\begin{assumption}\label{ass:M2}
	There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq W
\]
\end{assumption}

\begin{assumption}\label{ass:REINFORCE}
There is a constant $V<\infty$ such that, for each policy $\pol$:
\[
	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq V
\]
\end{assumption}

\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth with Lipschitz constant:
\[
	L = \frac{R(G^2+F)}{1-\gamma}
\]
\end{restatable}

@@

	\\

@@

		\nonumber\\%

@@

			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \\

@@

			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\

@@

	+G^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \\

@@

	+G^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\

@@

	+G^2\frac{R^2W}{(1-\gamma)^2B},

@@

where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.

@@

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}
\end{align}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
	%
	&\leq
	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
	%
	&\leq	
	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
	%
	&=
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
	%
	&\leq
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

>>>>

%%Theorems
\usepackage{amsthm}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{conj}{Conjecture}
\newtheorem{example}{Example}
%Restatable
\usepackage{thmtools, thm-restate}
\declaretheorem[numberwithin=section]{thm}
\declaretheorem[sibling=thm]{lemma}
\declaretheorem[sibling=thm]{corollary}
\declaretheorem[sibling=thm]{assumption}
\declaretheorem[sibling=thm]{theorem}
%e.g. ...
\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
% graphics
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
%% TODOs
% \setlength{\marginparwidth}{2cm}
\usepackage[colorinlistoftodos, textwidth=20mm]{todonotes}
\definecolor{citrine}{rgb}{0.89, 0.82, 0.04}
\definecolor{blued}{RGB}{70,197,221}
%%todo by Marcello
\newcommand{\todomarc}[1]{\todo[color=green, inline]{\small #1}}
\newcommand{\todomarcout}[1]{\todo[color=green]{\scriptsize #1}}
%%todo by Matteo Papini
\newcommand{\todomat}[1]{\todo[color=citrine, inline]{\small #1}}
\newcommand{\todomatout}[1]{\todo[color=citrine]{\scriptsize #1}}
%%todo by Matteo Pirotta
\newcommand{\todopir}[1]{\todo[color=blued, inline]{\small #1}}
\newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
%%todo by Damiano
\newcommand{\tododam}[1]{\todo[color=orange, inline]{\small #1}}
\newcommand{\tododamout}[1]{\todo[color=orange]{\scriptsize #1}}
%%todo by Giuseppe
\newcommand{\todobep}[1]{\todo[color=purple, inline]{\small #1}}
\newcommand{\todobepout}[1]{\todo[color=purple]{\scriptsize #1}}
%%%%%%

%%%CUSTOM COMMANDS%%%
%Math
\newcommand{\realspace}{\mathbb R}      % realspace
\newcommand{\transpose}[1]{{#1}^\texttt{T}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\EV}{\mathbb{E}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator*{\Cov}{\mathbb{C}ov}
\newcommand{\EVV}[2][\ppvect \in \ppspace]{\EV_{#1}\left[{#2}\right]}
\newcommand{\norm}[2][\infty]{\left\|#2\right\|_{#1}}
\newcommand{\Dij}[2]{\frac{\partial^{2}{#1}}{\partial{#2}_i\partial{#2}_j}}
\newcommand{\de}{\,\mathrm{d}}
\newcommand{\dotprod}[2]{\left\langle#1,#2\right\rangle}
%RL
\newcommand{\vtheta}{\boldsymbol{\theta}}
\newcommand{\Aspace}{\mathcal{A}}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\Tspace}{\mathcal{T}}
\newcommand{\Transition}{\mathcal{P}}
\newcommand{\Reward}{\mathcal{R}}
\newcommand{\stationary}{d_{\rho}^{\pi_{\vtheta}}(s)}
\newcommand{\policy}{\pi_{\vtheta}(a \mid s)}
\newcommand{\pol}{\pi_{\vtheta}}
\newcommand{\trajdistr}{\pi_{\vtheta}(\tau)}
\newcommand{\score}[2]{\nabla\log\pi_{#1}(#2)}
\newcommand{\Qfun}{Q^{\pi_{\vtheta}}(s,a)}
\newcommand{\Vfun}{V^{\pi_{\vtheta}}(s)}
\newcommand{\vTheta}{\boldsymbol{\Theta}}
\newcommand{\vphi}{\boldsymbol{\phi}}
\newcommand{\gradJ}[1]{\nabla J(#1)}
\newcommand{\gradApp}[2]{\hat{\nabla}_{#2}J(#1)}
\newcommand{\eqdef}{\mathrel{\mathop:}=}
\newcommand{\Dataset}{\mathcal{D}}
%Specific
\newcommand{\Ets}[2][t]{\mathbb{E}_{#1\mid s}\left[#2\right]}
\newcommand{\Covts}[3][t]{{\mathbb{C}\text{ov}}_{#1\mid s}\left(#2,#3\right)}
\newcommand{\Varts}[2][t]{{\mathbb{V}\text{ar}}_{#1\mid s}\left[#2\right]}
\newcommand{\gradBlack}[1]{\blacktriangledown J(#1)}
\newcommand{\gradIdeal}[1]{\overline{\blacktriangledown} J(#1)}
\newcommand{\VARRF}{V}
\newcommand{\GRADLOG}{G}
\newcommand{\VARIS}{W}
\newcommand{\HESSLOG}{F}
%%%%%%

@@

\section{Algorithm}
\todomat{Add minibatch}
In on-line policy gradient problems, stochastic gradient ascent (SGA) is not really a choice, but is dictated by the necessity of interacting with an unknown environment. We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{itemize}
	\item The objective $J(\vtheta)$ is typically non-convex;
	\item The dataset $\mathcal{T}$ can be infinite;
	\item The value of the parameter $\vtheta$ influences the sampling of trajectories;
\end{itemize}
To deal with non-convexity, we require $J(\vtheta)$ to be L-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and Softmax [appendix?]. Because of the infinite dataset, we can only rely on an estimate of the full gradient. Finally, we employ importance weighting \cite{rubinstein1981simulation} \cite{precup2000eligibility} to guarantee proper sampling of trajectories. We focus on the REINFORCE gradient estimator for simplicity, but the extension to G(PO)MDP is straightforward. The proposed SVRG-like gradient estimate, which we call Stochastic Variance Reduced Policy Gradient (SVRPG), is the following:
\begin{align*}
	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \mid \tilde{\vtheta})\Reward(\tau_j) \\
		&+ \nabla\log\pi(\tau_i \mid \vtheta)\Reward(\tau_i) 
		- \omega(\tau_i)\nabla\log\pi(\tau_i \mid \tilde{\vtheta})\Reward(\tau_i),
\end{align*}
where trajectories $\tau_j$ for $j=1,\dots,N$ are sampled from a snapshot policy $\pi_{\tilde{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pol$, and $\omega(\tau) = \frac{\pi(\tau\mid\tilde{\vtheta})}{\pi(\tau\mid\vtheta)}$ is an importance weight from $\pol$ to the snapshot policy $\pi_{\tilde{\vtheta}}$. We immediately have the following (proof in Appendix \ref{app:proofs}):

\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
The SVRPG gradient estimator is unbiased:
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
\end{restatable}
\todomat{Add lemma on zero variance at convergence}
The resulting SVRPG policy optimization method is detailed in Algorithm \ref{alg:svrpg}.
To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient and the step size $\alpha$. Adaptive selection of these meta-parameters is also possible. However, the next section will focus on providing converge guarantees for the plain SVRPG algorithm.

@@

\section{Convergence Guarantees}
\begin{assumption}\label{ass:bounded_score}
For each trajectory $\tau$, value of $\vtheta$ and dimensions $i,j$ there are positive constants $G,F<\infty$ such that:
\begin{align*}
	&\left|\nabla_{\theta_i}\log\pi_{\vtheta}(\tau)\right| \leq \GRADLOG \\
	&\left|\frac{\partial^2}{\partial\theta_i\partial\theta_j}\log\pi_{\vtheta}(\tau)\right| \leq \HESSLOG
\end{align*}
\end{assumption}

\begin{assumption}\label{ass:M2}
	There is a constant $W<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq \VARIS
\]
\end{assumption}

\begin{assumption}\label{ass:REINFORCE}
There is a constant $V<\infty$ such that, for each policy $\pol$:
\[
	\Tr\mathbb{V}ar\left(\gradApp{\vtheta}{1}\right) \leq \VARRF
\]
\end{assumption}

\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth.
\end{restatable}

@@

@@

		\label{eq:1}\\%

@@

			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\

@@

			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \label{eq:3}

@@

	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2}\frac{1}{B^2}\sum_{i=0}^{B-1}\Ets{(\omega(\tau_i)-1)^2} \label{eq:4}\\

@@

	+\GRADLOG^2\frac{R^2}{(1-\gamma)^2b^2}\sum_{i=0}^{B-1}\Varts{\omega(\tau_i)} \nonumber\\

@@

	+\GRADLOG^2\frac{R^2\VARRF}{(1-\gamma)^2B}, \label{eq:5}

@@

where (\ref{eq:1}) is from the definition of $\gradApp{\vtheta}{N}$, (\ref{eq:2}) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (\ref{eq:3}) is from L-smoothness, 
(\ref{eq:4}) is from Assumption \ref{ass:bounded_score}, and (\ref{eq:5}) is from Assumption \ref{ass:M2}.

@@

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:6}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \label{eq:7}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \label{eq:8}
\end{align}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
	%
	&\leq
	\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
	%
	&\leq	
	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\label{eq:9}\\
	%
	&=
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \label{eq:10}\\
	%
	&\leq
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} +\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{b\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(\frac{3}{2}+\alpha_t\beta_t+\alpha_t^2\right) + \alpha_t\frac{L^2}{b}+\frac{\alpha_t^2L^3}{2}, \\
	&\Psi_t = \alpha_t\left(1-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_t^2L+2\alpha_t^2c_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{G^2R^2(L+2c_{t+1})}{2(1-\gamma)^2}.
\end{align*}
In particular, the following constraints on $\alpha_t$ and $\beta_t$ are sufficient:
\begin{align*}
&0\leq\alpha_t < \frac{2(1-\nicefrac{c_{t+1}}{\beta_t})}{L+2c_{t+1}} \\
&\beta_t > c_{t+1}.
\end{align*}
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \label{eq:11}\\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \label{eq:12}\\
	&\geq
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}}^2 - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{\alpha_tL^2}{2}\Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}, \label{eq:13}
\end{align}
where (\ref{eq:11}) is from the L-smoothness of $J(\vtheta)$, (\ref{eq:12}) is from the SVRPG update, and (\ref{eq:13}) is from Lemma \ref{lemma:aux0}.
